---
title: "IST 687 Group Project"
date: '2022-11-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Install and Library Needed Packages 
```{r,echo = T, results= 'hide', error=FALSE, warning =FALSE, message=FALSE}
#Library Needed Packages
library(tidyverse)
library(ggplot2)
library(caret)
library(kernlab)
library(imputeTS)
library(e1071)
library(rpart)
library(rpart.plot)
```

2. Load Dataset 
```{r}
data <- read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/HMO_data.csv")
```

3. Change Character Variables to Factor Variables for Analysis 
```{r}
str(data)
data$smoker = as.factor(data$smoker)
data$location = as.factor(data$location)
data$location_type = as.factor(data$location_type)
data$education_level = as.factor(data$education_level)
data$yearly_physical = as.factor(data$yearly_physical)
data$exercise = as.factor(data$exercise)
data$married = as.factor(data$married)
data$hypertension = as.factor(data$hypertension)
data$gender = as.factor(data$gender)
```

4. Fix NA values and remove ones that can't be interpolated 
```{r}
sum(is.na(data)) #158 Null Values
data = na_interpolation(data)
sum(is.na(data)) #80 Null Values remain. Can't be changed because character value
data = drop_na(data) #Remove these values 
sum(is.na(data)) #0 Null values left
 
```

5. Linear Model to see which variables are predictive of Cost 
```{r}
lm <- lm(cost ~., data = data)
summary(lm)

#Age, BMI, Children, Being a Smoker, Not Exercising, living in New York, and having Hypertension raise costs of healthcare
```

6. Create variable for Expensive Healthcare
```{r}
data$expensive[data$cost >= 5000] <-  "Expensive" 
#top 25% of costs are labeled expensive 
data$expensive[is.na(data$expensive)] <-  "Normal"
data$expensive = as.factor(data$expensive) #change to factor variable
```

7. Partition data into training and test set 
```{r}
new_data = data [-14] #removing "cost" variable from data. 
trainList <- createDataPartition(y=new_data$expensive,p=.4,list=FALSE) #40% train data
train <- new_data[trainList,]
test <- new_data[-trainList,]
```

8. Create SVM Classification Model 
```{r}
model <- ksvm(data = train, expensive ~., C=5,CV=3, prob.model = TRUE )
model

svmPred <- predict(model, newdata = test) #predict using test data
```

9. Look at Accuracy of Model 
```{r}
confusionMatrix(svmPred, test$expensive) 

#87.8% accurate with 59.6% sensitivity 
```
10. Rpart Model 
```{r}
rpartmodel <- rpart(expensive ~ ., data = train)
rpart.plot(rpartmodel)

rpartmodel <- rpart(expensive ~ ., data = test)
predictValues <- predict(rpartmodel, type = "class")
table(predictValues)

```
11. Look at Rpart Accuracy 
```{r}
confusionMatrix(predictValues, test$expensive)

#89.0% accuracy with 61.1% sensitivity. 
#This is the better model

```

12. Visualizations 
```{r}
ggplot(data, aes(expensive, age)) + geom_boxplot() + ggtitle("Healthcare Cost by Age")
ggplot(data, aes(expensive, bmi)) + geom_boxplot() + ggtitle("Healtchare Cost by BMI")
ggplot(data, aes(expensive, children)) + geom_boxplot() + ggtitle("Healthcare Cost by number of Children")
```

13. Subset Data into Expensive and Normal 
```{r}
expensive = filter(data, expensive == "Expensive")
normal = filter(data, expensive == "Normal")
data$smokeryes <-ifelse(data$smoker == 'yes', 1,0)
data$smokerno <- ifelse(data$smoker == 'no',1,0)

#there's more no. of non smokers but smokers clearly have a higher cost the graph is right skewed
ggplot(expensive, aes(x = cost)) +
geom_histogram(aes(color = smoker), fill = "white",position = "identity", bins = 30) +
scale_color_manual(values = c("#00AFBB", "#E7B800")) +geom_vline(xintercept = mean(expensive$cost)) +geom_vline(xintercept = mean(normal$cost)) + ggtitle("Cost of Healthcare Smokers vs Non-Smokers")


#looks like being a smoking has no effect on BMI 
ggplot(data, aes(x = bmi)) +
geom_histogram(aes(color =smoker), fill = "white",position = "identity", bins = 30) +
scale_color_manual(values = c("#00AFBB", "#E7B800")) #+geom_vline(xintercept = data(expensive$cost))

#linear relationship between age and cost for normal dataset
ggplot(normal, aes(x=age, y=cost, color=cost))+geom_point()

#non linear relationship. if this is an exponential trend, can we show a trend line??
ggplot(expensive, aes(x=age, y=cost, color=cost))+geom_point() # ylim(5000,65000)

plot(data$smoker)
plot(data$hypertension)

```


14. US Map plotting for costs for each state in US
```{r}
library(ggplot2)
library(maps)
library(ggmap)
newDF <- data %>% group_by(location) %>% summarise(avgcost = sum(cost))
#Load the pre-defined dataset 'state' is loaded in us dataframe
us<- map_data("state")
#Change state_name column values to lowercase 
us$state_name <- tolower(us$region)
#check the structure of the us dataframe
str(us)
newDF$location <- tolower(newDF$location)
str(newDF)
#merge() us and dfSimple dataframe, the merge is done based on state_name
#column in both dataframes
farewithgeom <- merge(us,newDF,by.y="location",by.x="state_name") 
#arrange the order of the popwithgeom dataframe
#structure of the popwithgeom
str(farewithgeom)
#Calculate the bounding box to define the us states
bb <- c(left = min(us$long), bottom = min(us$lat),right = max(us$long), top = max(us$lat))
map <- get_stamenmap(bbox = bb, zoom=5)
#plot map using ggmap and add the color shading based on the Pop
#column of dfNew dataframe
ggmap(map) + geom_polygon(data=farewithgeom,color="black", alpha=0.9,aes(x=long,y=lat,group=group,fill=avgcost))
```

15. 
```{r}
#save the model and the data
test %>% mutate_if(is.character, utf8::utf8_encode)
save(rpartmodel, file = "rpart.rda")
save(test, file = "testdata.rda")

```


data(testdata.rda)
```{r}

```
